<!-- badges: start -->
[![R build status](https://github.com/ropenscilabs/statistical-software-review/workflows/R-CMD-check/badge.svg)](https://github.com/ropenscilabs/statistical-software-review/actions)
<!-- badges: end -->


<!-- README.md is generated from README.Rmd. Please edit that file -->

# Statistical Software Review

This repository is primarily intended for submissions to rOpenSci's expanded
system of software peer review to encompass statistical software. The documents
of the project, including standards and procedures which will be used to guide
the process, are in the [statistical-software-review-book
repository](https://github.com/ropenscilabs/statistical-software-review-book).

## But it appears to be an R package?

Ah, that's true. The repository also contains a few functions both for the use
of developers submitting packages for review, and for reviewers. This
repository also contains an R package called `statsoftrev` which developers and
reviewers will need to install with,

```{r install, eval = FALSE}
# install.packages("remotes")
remotes::install_github("ropenscilabs/statistical-software-review")
```

It can then be loaded for use with,
```{r lib}
library (statsoftrev)
```

All functions of the packages are prefixed with `rssr_`, for "**r**OpenSci
**S**tatistical **S**oftware **R**eview." One function provides a list of
currently developed categories of statistical software for which standards have
been developed, along with links to the online standards for each category:

```{r available}
rssr_available_categories ()
```

Any software within one or more of these categories may be considered for
review. The remaining functions are intended for submitting authors, reviewers,
and rOpenSci editors, and serve the two primary tasks of:

1. Obtaining and processing checklists of standards; and 
2. Enabling and processing [`roxygen2`](https://roxygen2.r-lib.org) tags to
   align package functions with standards.

These functions will be needed, for different purposes, by both developers and
reviewers. The book on [Statistical Software
Review](https://ropenscilabs.github.io/statistical-software-review-book/index.html)
details the procedures to be followed by both developers and reviewers to align
and asses software against the standards. This package enables copies of those
standards to be obtained as markdown-formatted checklists, as described in the
first of the following sections. Developers are also expected to align the
various functions of their software with specific standards using
project-specific [`roxygen2`](https://roxygen2.r-lib.org) tags, as described in
the subsequent section.

## 1. Checklists of standards

Both developers and reviewers will need to obtain checklists of general and
category-specific standards relevant to the package, with detailed procedures
given in the main [project
book](https://ropenscilabs.github.io/statistical-software-review-book/index.html).
The function
[`rssr_standards_checklist()`](https://ropenscilabs.github.io/statistical-software-review/reference/rssr_standards_checklist.html)
converts one or more of the currently written standards into
a rmarkdown-formatted checklist. These checklists are also copied directly to
a local clipboard, ready to be pasted directly into a local file or a github
issue. This function always extracts the checklist of [General
Standards](https://ropenscilabs.github.io/statistical-software-review-book/standards.html#general-standards-for-statistical-software),
along with any specified categories appended as additional checklists.

```{r checklist-fakey, eval = FALSE}
x <- rssr_standards_checklist (category = c ("regression", "ml"))
```
```{r checklist, eval = TRUE, echo = FALSE}
Sys.setenv ("CLIPR_ALLOW" = "TRUE")
x <- rssr_standards_checklist (category = c ("regression", "ml"))
```

The resultant object is a character vector which starts like this:

```{r checklist-out}
head (x)
```

The names of the categories are those given in the "category" column of the
output of
[`rssr_available_categories()`](https://ropenscilabs.github.io/statistical-software-review/reference/rssr_available_categories.html)
shown above. Calling that function will give you a copy of all relevant
standards for the specified category (or categories) in your clipboard which
you can paste anywhere. The function also has a `filename` argument which can
be used to save the checklist directly to a local file.

### 1.1 Fillling out checklists

While it is possible to paste a checklist directly into a github issue and
complete it there, we recommend that you first save the checklist to a local
file (using the `file` option of the
[`rssr_standards_checklist()`](https://ropenscilabs.github.io/statistical-software-review/reference/rssr_standards_checklist.html)
function) and use that local version to complete your assessment. Each standard
which you judge the software to meet can be checked by replacing the space
between the open square brackets with an "x". Whether or not the checkbox is
filled, please replace the text for each standard with a comment explaining how
the software does or does not fulfil that standard. Please do not remove or
change the identifiers for each standard.

### 1.2 Inapplicable Standards

Any standards deemed inapplicable to a given software package should be
checked, the text potentially modified to indicate how or why those standards
do not apply, and the text then appended with “N/A” in bold face (achieved by
surrounding with either two asterisks or two underscores, “\*\*N/A\*\*” or
“\_\_N/A\_\_”). Multiple inapplicable standards may be combined in a single
statement through separating by a double dash ("\-\-"), as in the following
example:

- [x] **G4.10**--**G4.12** It is appropriate for this software not to implement
      any extended tests **N/A**

That example demonstrates that standards deemed to be inapplicable to a given
software package should be considered fulfilled and checked appropriately.
There may of course be standards which are deemed inapplicable, yet which you
as a reviewer judge should actually be applied. For example, a package may not
implement extended tests, yet may have several functions which take too long to
run to be tested within a normal test suite. Such a package should have an
extended test suite, and so the relevant general standards **G4.10**--**G4.12**
should be considered applicable. In such cases, please simply leave these
standards unchecked, and alter the relevant text to clarify why you think those
standards should apply.

### 1.3 Check the formatting of your checklist

Having completed your checklist, please use one final function,
[`rssr_checklist_check()`](https://ropenscilabs.github.io/statistical-software-review/reference/rssr_checklist_check.html),
to ensure that common formatting mistakes can be rectified. This is important
to facilitate our automated analyses of checklists and software assessments,
through ensuring that all items are in standardised formats. Simply pass the
name of the local file containing your checklist to this function, and the file
will be updated with any necessary changes, as well as copied again to your
local clipboard ready to be pasted into the github review issue.

## 2. `roxygen2` tags

The `statsoftrev` package associated with this repository includes functions to
process project-specific [`roxygen2`](https://roxygen2.r-lib.org) tags of
`#' @rssr`, to align package functions with specific standards. Packages can
enable these tags by adding or modifying the `Roxygen` line of a package's
`DESCRIPTION` file to include the `statsoftrev::rssr_roclet`. The result might
look something like the following:

```
Roxygen: list(markdown = TRUE, roclets = c("namespace", "rd", "statsoftref::rssr_roclet"))
```

You do not need to import or in any way depend on the `statsoftrev` package,
but merely add the "roclet" as in the above line. For developers, this will
have the sole effect of enabling [`roxygen2`](https://roxygen2.r-lib.org) to
ignore all `@rssr` tags in your documentation, and to proceed to generate
documentation as normal. (Documentation will still be generated even without
adding the "roclet" to your `DESCRIPTION` file, but you'll see warnings about
"@rssr unknown tag".)

Developers are required to align the functions of their software with specific
standards which those functions address by inserting `@rssr` tags in their
function documentation, Each standard should be tagged at least once within
a package by a `#' @rssr <standard-number>` tag, where `<standard number>`
refers to one or more of the general or category-specific standards.

For example, the first standard illustrated above is,
```{r g1, echo = FALSE}
print (grep ("G1\\.0", x, value = TRUE))
```

Such primary references are commonly provided in the `DESCRIPTION` file, but
should also be documented within the code itself. One common place for
references is in the primary package documentation, typically within a file
named, `R/<mypkg>-package.R`, documented as described in the [`roxygen2`
documentation](https://roxygen2.r-lib.org/articles/rd.html#packages-1).
Packages initially generated with [`usethis`](https://usethis.r-lib.org) with
automatically have this package-level documentation. Primary references may
then be inserted anywhere within this package-level documentation, typically
with a [`\doi{}`
tag](https://cran.r-project.org/doc/manuals/R-exts.html#User_002ddefined-macros).
The documentation of that primary reference can be aligned to [Standard
**G1.0**](https://ropenscilabs.github.io/statistical-software-review-book/standards.html#documentation)
by adding the following [`roxygen2`](https://roxygen2.r-lib.org) tag:

```{r rssr-tag}
#' @rssr G1.0
```
A full example might look something like this:
```{r rssr-tag-full, eval = FALSE}
#' @details
#' This package is an implementation of the algorithm described in
#' \doi{<reference-to-published-algorithm>}.
#' @keywords internal
#' @rssr G1.0
"_PACKAGE"
```

These `@rssr` tags do not appear in any resultant documentation, rather they
only serve to enable identification of which components of a package address
which standards.
