<!-- badges: start -->
[![R build status](https://github.com/ropenscilabs/statistical-software-review/workflows/R-CMD-check/badge.svg)](https://github.com/ropenscilabs/statistical-software-review/actions)
<!-- badges: end -->


<!-- README.md is generated from README.Rmd. Please edit that file -->

# Statistical Software Review

This repository is primarily intended for submissions to rOpenSci's expanded system of
software peer review to encompass statistical software. The documents of the project, including standards and procedures which will be used to guide the process, are in the [statistical-software-review-book repository](https://github.com/ropenscilabs/statistical-software-review-book).

## But it appears to be an R package?

Ah, that's true. The repository also contains a few functions intended to aid the review process, bundled within an R package called `statsoftrev`. The functions themselves are prefixed with `rssr_` for **r**OpenSci **s**tatistical **s**oftware **r**eview. At present these functions are

```{r fn-list}
library (statsoftrev)
ls ("package:statsoftrev")
```

The first of these functions provides a list of currently developed categories of statistical software for which standards have been developed.

```{r available}
rssr_available_categories ()
```

The second function converts one or more of the currently written standards into a rmarkdown-formatted checklist for the benefit of reviewers. These checklists are also copied directly to a local clipboard, ready to be pasted directly into a local file or a github issue. This function always extracts the checklist of [General Standards](https://ropenscilabs.github.io/statistical-software-review-book/standards.html#general-standards-for-statistical-software), with any specified categories appended as additional checklists.

```{r checklist-fakey, eval = FALSE}
x <- rssr_standards_checklist (category = "regression")
```
```{r checklist, eval = TRUE, echo = FALSE}
Sys.setenv ("CLIPR_ALLOW" = "TRUE")
x <- rssr_standards_checklist (category = "regression")
```

The resultant object is a character vector which starts like this:

```{r checklist-out}
head (x)
```



